---
title: "Population-level muscle synergy analysis with actual data"
author: "Yusuke Matsui"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

This RMarkdown document describes a comprehensive synergy analysis pipeline using NMF. It includes data preprocessing, individual synergy analysis, synergy matching, population synergy analysis, and visualization.

## Load Libraries and Data

```{r setup, include=TRUE, eval=FALSE}
# Load required libraries
library(parallel)
library(data.table)
library(NMF)
library(tidyverse)

# Load custom functions and data
source("simulation_function.R")
source("modules.R")
load("m_arr_lt.rda") # Load muscle activity data
```

## Define Parameters and Preprocessing

```{r parameters, eval=FALSE}
# Define muscle names and other parameters
mn <- c("Rt_TA", "Lt_TA", "Rt_LG", "Lt_LG", "Rt_MG", "Lt_MG", 
        "Rt_VM", "Lt_VM", "Rt_Ham", "Lt_Ham")
nsyn <- 4  # Number of synergies
nrun <- 10 # Number of NMF runs
set.seed(1234) # For reproducibility

# Preprocess data: Replace negative values with zero
simdat <- m_arr_lt
simdat[simdat < 0] <- 0
```

## Visualize Synergy Comparisons

```{r visualize-comparisons, fig.width=10, fig.height=10, eval=FALSE}
# Visualize concatenated, individual, and population synergies
library(ggplot2)
library(gridExtra)

# Prepare H data for visualization
Hs_mean <- t(sapply(1:nsyn, function(ii) rowMeans(sapply(Hs, function(xx) xx[ii, ]))))
Hs_mean_rel <- 100 * Hs_mean / max(Hs_mean)
H_pop_rel <- 100 * H_pop / max(H_pop)

df_individual_H <- data.frame(Hs_mean_rel)
colnames(df_individual_H) <- mn
df_individual_H$Synergy <- paste("Synergy", 1:nsyn)
df_individual_H_long <- tidyr::gather(df_individual_H, key = "Muscle", value = "Value", -Synergy)

df_population_H <- data.frame(H_pop_rel)
colnames(df_population_H) <- mn
df_population_H$Synergy <- paste("Synergy", 1:nsyn)
df_population_H_long <- tidyr::gather(df_population_H, key = "Muscle", value = "Value", -Synergy)

df_individual_H_long$Method <- "Individual"
df_population_H_long$Method <- "Population"

df_combined_H <- rbind(df_individual_H_long, df_population_H_long)

# Plot H matrix
gg_H <- ggplot(data = df_combined_H, aes(x = Muscle, y = Value, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(Synergy ~ Method, nrow = 4, ncol = 3) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5)) +
  ylab("Value [%]") +
  xlab("Muscle") +
  ggtitle("H Matrix Synergy Comparison")
print(gg_H)

# Prepare W data for visualization
W_pop_rel <- apply(W_pop, 2, function(x) 100 * x / max(x))
W_concat_mean <- apply(W_concat, c(1, 3), mean)
W_concat_rel <- apply(W_concat_mean, 2, function(x) 100 * x / max(x))

df_population_W <- data.frame(W_pop_rel)
colnames(df_population_W) <- paste("Synergy", 1:nsyn)
df_population_W$Time <- seq_len(nrow(W_pop_rel))
df_population_W_long <- tidyr::gather(df_population_W, key = "Synergy", value = "Value", -Time)

# Plot W matrix
gg_W <- ggplot(data = df_population_W_long, aes(x = Time, y = Value, color = Synergy)) +
  geom_line(size = 1.2) +
  ylab("Value [%]") +
  xlab("Time") +
  ggtitle("W Matrix Synergy Comparison")
print(gg_W)
```

## Save Results and Visualizations

```{r save-results, eval=FALSE}
# Save all results to the current directory
save.image("results.rda")

# Save the visualizations
ggsave("H_matrix_synergy_comparison.pdf", plot = gg_H, width = 10, height = 10)
ggsave("W_matrix_synergy_comparison.pdf", plot = gg_W, width = 10, height = 10)
```

# Conclusion

This document provides a detailed pipeline for synergy analysis using NMF, including individual and population synergy calculations, synergy matching, and visualization of both `H` and `W` matrices.
