---
title: "Population-level muscle synergy analysis with outliers"
author: "Yusuke Matsui"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

This document details a comprehensive analysis pipeline for synergy simulations, incorporating simulation generation, individual synergy computation, population synergy computation, and visualization of results.

## Load Libraries and Data

```{r setup, include=TRUE, eval=FALSE}
library(parallel)
library(data.table)
library(NMF)
library(tidyverse)

# Load custom functions and references
source("simulation_function.R")
source("modules.R")
ref <- readRDS("m_arr_pc_mean_lt.rds")
```

## Parameter Initialization

```{r initialize-parameters, eval=FALSE}
param <- seq(10, 100, by=10)  # Sample sizes
mn <- c("Rt_TA", "Lt_TA", "Rt_LG", "Lt_LG", "Rt_MG", "Lt_MG", "Rt_VM", "Lt_VM", "Rt_Ham", "Lt_Ham")
nsyn <- 4  # Number of synergies
nrun <- 10  # Number of NMF runs
set.seed(1234)
```

## Generate Simulation Data

```{r generate-simulation-data, eval=FALSE}
for (s in seq_along(param)) {
  n <- param[s]
  ns <- ceiling(c(0.9, 0.1) * n)  # Define mixture ratio
  st <- c(1, cumsum(ns[-length(ns)]) + 1)
  ed <- cumsum(ns)
  simdat <- array(NA, dim = c(nrow(ref), ncol(ref), n))
  
  for (i in seq_along(ns)) {
    for (j in seq_len(ncol(ref))) {
      set.seed(100 * i + 50 * j)
      if (i == 1) {
        simdat[, j, st[i]:ed[i]] <- simEMG_envelope_multi(ref[, j], nsim = ns[i],
                                                          n_jump = 0, amp_rate = rbeta(1, 3, 1),
                                                          noise_rate = 1, plot.it = FALSE)
      } else {
        simdat[, j, st[i]:ed[i]] <- simEMG_envelope_multi(ref[, j], nsim = ns[i],
                                                          n_jump = 2, amp_rate = 1,
                                                          noise_rate = 1.5, plot.it = FALSE)
      }
    }
  }
  simdat[simdat < 0] <- 0
}
```

## Compute Individual Synergies

```{r individual-synergies, eval=FALSE}
cl <- makeCluster(10)
clusterExport(cl, c("simdat", "nsyn", "nrun"))

ind_syn <- parLapply(cl, 1:dim(simdat)[3], function(ii) {
  library(NMF)
  nmf(simdat[, , ii], "brunet", rank = nsyn, seed = "nndsvd", nrun = nrun)
})
stopCluster(cl)
```

## Perform Synergy Matching

```{r synergy-matching, eval=FALSE}
synMatch <- matrix(NA, nrow = length(ind_syn), ncol = nsyn)
for (i in 1:nrow(synMatch)) {
  synCor <- apply(ind_syn[[1]]@fit@H, 1, function(xx) apply(ind_syn[[i]]@fit@H, 1, function(yy) cov(xx, yy)))
  comb <- apply(synCor, 2, function(x) which(x > 0), simplify = FALSE)
  select_comb <- comb[apply(comb, 1, function(x) length(unique(x)) == nsyn), ]
  if (nrow(select_comb) > 0) {
    cost <- apply(select_comb, 1, function(x) sum(sapply(seq_along(x), function(y) synCor[x[y], y])))
    synMatch[i, ] <- unlist(select_comb[which.max(cost), ])
  }
}
```

## Compute Population Synergies

```{r population-synergies, eval=FALSE}
simdat_mu <- apply(simdat, c(1, 2), mean)
simdat_mu[simdat_mu < 0] <- 0
nmf_pop <- nmf(simdat_mu, "brunet", rank = nsyn, seed = "nndsvd", nrun = nrun)
W_pop <- nmf_pop@fit@W
H_pop <- nmf_pop@fit@H
```

## Visualize Results

### Visualize W Matrices

```{r visualize-W, eval=FALSE}
library(ggplot2)

W_pop_rel <- apply(W_pop, 2, function(x) 100 * x / max(x))
df_W <- data.frame(W_pop_rel)
colnames(df_W) <- paste("Synergy", 1:nsyn)
df_W$Time <- seq_len(nrow(W_pop_rel))
df_W_long <- tidyr::gather(df_W, key = "Synergy", value = "Value", -Time)

ggplot(data = df_W_long, aes(x = Time, y = Value, color = Synergy)) +
  geom_line(size = 1.2) +
  labs(x = "Time", y = "Value [%]", title = "W Matrix Synergy Comparison")
```

### Visualize H Matrices

```{r visualize-H, eval=FALSE}
H_pop_rel <- t(apply(H_pop, 1, function(x) 100 * x / max(x)))
df_H <- data.frame(H_pop_rel)
colnames(df_H) <- mn
df_H$Synergy <- paste("Synergy", 1:nsyn)
df_H_long <- tidyr::gather(df_H, key = "Muscle", value = "Value", -Synergy)

ggplot(data = df_H_long, aes(x = Muscle, y = Value, fill = Synergy)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Muscle", y = "Value [%]", title = "H Matrix Synergy Comparison")
```

## Save Results

```{r save-results, eval=FALSE}
save.image(file = paste0("results_", Sys.Date(), ".rda"))
```

# Conclusion

This document provides a comprehensive pipeline for generating, analyzing, and visualizing muscle synergies. The approach covers simulation, synergy computation, matching, and population-level analysis.
