---
title: "Population-level muscle synergy analysis with different sample sizes"
author: "Yusuke Matsui"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

This RMarkdown document provides a comprehensive pipeline for synergy analysis using NMF, including visualization of both `H` and `W` matrices for individual, concatenated, and population synergies.

## Load Libraries and Data

```{r setup, include=TRUE,eval=FALSE}
# Load required libraries
library(parallel)
library(data.table)
library(NMF)
library(tidyverse)

# Load custom functions and data
source("simulation_function.R")
source("modules.R")
load("m_arr_lt.rda") # Load muscle activity data
```

## Define Parameters and Preprocessing

```{r parameters,eval=FALSE}
# Define muscle names and other parameters
mn <- c("Rt_TA", "Lt_TA", "Rt_LG", "Lt_LG", "Rt_MG", "Lt_MG", 
        "Rt_VM", "Lt_VM", "Rt_Ham", "Lt_Ham")
nsyn <- 4  # Number of synergies
nrun <- 10 # Number of NMF runs
set.seed(1234) # For reproducibility

# Preprocess data: Replace negative values with zero
simdat <- m_arr_lt
simdat[simdat < 0] <- 0
```

## Compute Individual Synergies

```{r individual-synergies,eval=FALSE}
# Set up parallel processing
cl <- makeCluster(7) # Use 7 cores
clusterExport(cl, c("simdat", "nsyn", "nrun"))

# Compute synergies for each individual using NMF
ind_syn <- parLapply(cl, 1:dim(simdat)[3], function(ii) {
  library(NMF)
  nmf(simdat[, , ii], "brunet", rank = nsyn, seed = "nndsvd", nrun = nrun)
})

# Stop the cluster
stopCluster(cl)
```

## Perform Synergy Matching

```{r synergy-matching,eval=FALSE}
# Initialize synergy matching matrix
refidx <- 1
synMatch <- matrix(NA, nrow = length(ind_syn), ncol = nsyn)
rownames(synMatch) <- paste("ind", 1:length(ind_syn))
colnames(synMatch) <- paste("synergy", 1:nsyn)

# Match synergies across individuals
for (i in 1:nrow(synMatch)) {
  synCor <- apply(ind_syn[[refidx]]@fit@H, 1, function(xx) 
    apply(ind_syn[[i]]@fit@H, 1, function(yy) cov(xx, yy)))
  
  comb <- apply(synCor, 2, function(x) which(x > 0), simplify = FALSE)
  len <- sapply(comb, length)
  
  if (all(len == 1)) {
    select_comb <- matrix(unlist(comb), nrow = 1)
  } else {
    comb <- do.call(expand.grid, comb)
    select_comb <- comb[apply(comb, 1, function(x) length(table(x)) == nsyn), ]
  }
  
  if (nrow(select_comb) > 0) {
    cost <- apply(select_comb, 1, function(x) 
      sum(sapply(seq_along(x), function(y) synCor[x[y], y])))
    idx <- which.max(cost)
    synMatch[i, ] <- unlist(select_comb[idx, ])
  }
}

# Remove invalid rows
rmidx <- apply(synMatch, 1, function(x) any(is.na(x)))
synMatch2 <- synMatch[!rmidx, ]
ind_syn2 <- ind_syn[!rmidx]
```

## Compute Population Synergies

```{r population-synergies,eval=FALSE}
# Compute mean muscle activity
simdat_mu <- apply(simdat, c(1, 2), mean)
simdat_mu[simdat_mu < 0] <- 0

# Perform NMF for population-level synergy
nmf_pop <- nmf(simdat_mu, "brunet", rank = nsyn, seed = "nndsvd", nrun = nrun)

# Extract population-level matrices
W_pop <- nmf_pop@fit@W
H_pop <- nmf_pop@fit@H
colnames(H_pop) <- mn
```

## Visualize `H` and `W` Matrices

### Visualize `H` Matrix

```{r visualize-h, fig.width=10, fig.height=6,eval=FALSE}
# Prepare data for visualization of H matrix
Hs_mean <- t(sapply(1:nsyn, function(ii) rowMeans(sapply(Hs, function(xx) xx[ii, ]))))
Hs_mean_rel <- 100 * Hs_mean / max(Hs_mean)
H_pop_rel <- 100 * H_pop / max(H_pop)

df_individual <- data.frame(Hs_mean_rel)
colnames(df_individual) <- mn
df_individual$Synergy <- paste("Synergy", 1:nsyn)
df_individual_long <- tidyr::gather(df_individual, key = "Muscle", value = "Value", -Synergy)

df_population <- data.frame(H_pop_rel)
colnames(df_population) <- mn
df_population$Synergy <- paste("Synergy", 1:nsyn)
df_population_long <- tidyr::gather(df_population, key = "Muscle", value = "Value", -Synergy)

# Combine data for visualization
df_individual_long$Method <- "Individual"
df_population_long$Method <- "Population"

df_combined <- rbind(df_individual_long, df_population_long)

# Plot H matrix comparisons
gg_h <- ggplot(data = df_combined, aes(x = Muscle, y = Value, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(Synergy ~ Method, nrow = 4, ncol = 3) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5)) +
  ylab("Value [%]") +
  xlab("Muscle") +
  ggtitle("H Matrix Synergy Comparison")
print(gg_h)
```

### Visualize `W` Matrix

```{r visualize-w, fig.width=10, fig.height=6,eval=FALSE}
# Prepare data for visualization of W matrix
W_pop_rel <- apply(W_pop, 2, function(x) 100 * x / max(x))
W_concat <- array(nmf_pop@fit@W, dim = c(dim(simdat)[1], dim(simdat)[3], nsyn))
W_concat_mean <- apply(W_concat, c(1, 3), mean)

df_W_pop <- data.frame(W_pop_rel)
colnames(df_W_pop) <- paste("Synergy", 1:nsyn)
df_W_pop$Time <- seq_len(nrow(W_pop_rel))
df_W_long <- tidyr::gather(df_W_pop, key = "Synergy", value = "Value", -Time)

# Plot W matrix comparisons
gg_w <- ggplot(data = df_W_long, aes(x = Time, y = Value, color = Synergy)) +
  geom_line(size = 1.2) +
  ylab("Value [%]") +
  xlab("Time") +
  ggtitle("W Matrix Synergy Comparison")
print(gg_w)
```

## Save Results and Visualizations

```{r save-results,eval=FALSE}
# Save all results to the current directory
save.image("results.rda")

# Save the visualizations
ggsave("H_matrix_synergy_comparison.pdf", plot = gg_h, width = 10, height = 10)
ggsave("W_matrix_synergy_comparison.pdf", plot = gg_w, width = 10, height = 10)
```

# Conclusion

This document provides a complete pipeline for synergy analysis using NMF, including individual and population synergy calculations, synergy matching, and visualization of both `H` and `W` matrices.
